\documentclass[10pt]{article}
 \usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb,amsfonts,color, titling}
 \usepackage{listings}

\setlength{\droptitle}{-20mm} 

\usepackage[colorlinks=true,linkcolor=red,urlcolor=blue]{hyperref}

\title{Pre-class assignment \#28}
\author{Brian O'Shea, \\PHY-905-003, Computational Astrophysics and
  Astrostatistics\\Spring 2017}
 \date{} % leave blank to have no date

\begin{document}
 
\maketitle

\vspace{-10mm}

\noindent
\textbf{This assignment is due the evening of Tuesday April 25, 2017.}
 Turn in all materials via GitHub.

\vspace{5mm}

\noindent
\textbf{Reading:}

\begin{enumerate}

\item Chapters 4, 5, and 11 of
  \href{https://e-reports-ext.llnl.gov/pdf/301321.pdf}{Radiation
    Hydrodynamics} by John Castor (available at that link and in the
  included PDF).  Chapter 4 covers radiation basics (which you may
  already know if you have taken an interstellar processes class);
  Chapter 5 covers steady-state radiative transfer; and Chapter 11
  surveys some of the numerical techniques used in radiation
  transport.  You will be randomly selected to implement either a
  flux-limited diffusion or Monte Carlo radiation transport method;
  pay particular attention to those.

\end{enumerate}

\noindent
\textbf{Your assignment:}  Answer the questions below in the file \texttt{ANSWERS.md}:

\begin{enumerate}

\item How does making the steady-state simplify the calculation of
  radiation transport?

\item What are the ``moments'' of the transport equations, and why do
  we care about that?

\item What are the relative advantages and disadvantages of the flux-limited
  diffusion  and the Monte Carlo approaches to radiation transport?

\item Why is it that scattering of radiation is such a challenge in
  numerical radiation transport?

\item What are any other questions that you have about the material
you read, any points that are not clear, or anything you'd like to
know more about.  Aim for at least 3 questions/unclear points/etc.

\end{enumerate}

\noindent \textbf{Handing it in:} Include 
your answers to the questions (in the file \texttt{ANSWERS.md})
in your assignment.

\end{document}
