\documentclass[10pt]{article}
 \usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb,amsfonts,color, titling}
 \usepackage{listings}

\setlength{\droptitle}{-20mm} 

\usepackage[colorlinks=true,linkcolor=red,urlcolor=blue]{hyperref}

\title{Pre-class assignment \# 11}
\author{Brian O'Shea, \\PHY-905-005, Computational Astrophysics and
  Astrostatistics\\Spring 2023}
 \date{} % leave blank to have no date

\begin{document}
 
\maketitle

\vspace{-5mm}

\noindent
\textbf{This assignment is due the evening of Wednesday Feb. 15, 2023.}
 Turn in all materials via GitHub.

\vspace{5mm}

\noindent
\textbf{Reading:}

\begin{enumerate}

\item Chapter 6 (``Burgers' Equation'') of Mike Zingale's
\href{http://bender.astro.sunysb.edu/hydro_by_example/CompHydroTutorial.pdf}{Computational
Hydrodynamics Tutorial}, which you can find in the pre-class
assignment for Day 8.  Focus on Sections 6.1-6.2, but make sure to
read the remainder of the chapter as well (which is only a few pages).

\item
\href{https://en.wikipedia.org/wiki/Burgers%27_equation}{Wikipedia
page on Burgers' Equation} (optional, but the animations are useful to
look at).

\end{enumerate}

\noindent
\textbf{Your assignment:}  


\begin{enumerate}
  
\item Extend your code from the previous class to implement the finite
  volume (i.e., conservative-form) solution to the inviscid Burgers'
  equation (Eqtn. 6.1 in Zingale) using whichever limiter you
  implemented in class and Dirichlet boundary conditions.  Set up the rarefaction wave initial
  conditions (shown in Figure 6.4 and described in the text) and the
  sinusoidal initial conditions (Equation 6.19 and Figure 6.5), and
  verify that they evolve correctly until the wave hits the boundary
  of the box.

\item What are at least two questions that you have from the readings
that you'd like us to address in class?
Put those questions in \texttt{ANSWERS.md}.

\end{enumerate}


\noindent \textbf{Handing it in:} Include your code, your plots, and
your answers to the questions about the readings (in the file \texttt{ANSWERS.md})
in your assignment.

\end{document}
