\documentclass[10pt]{article}
 \usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb,amsfonts,color, titling}
 \usepackage{listings}

\setlength{\droptitle}{-20mm} 

\usepackage[colorlinks=true,linkcolor=red,urlcolor=blue]{hyperref}

\title{Pre-class assignment \#13}
\author{PHY-905-005\\Computational Astrophysics and
  Astrostatistics\\Spring 2023}
 \date{} % leave blank to have no date

\begin{document}
 
\maketitle

\vspace{-5mm}

\noindent
\textbf{This assignment is due the evening of Wednesday March 15, 2023.}
 Turn in all materials via GitHub.

\vspace{5mm}

\noindent
\textbf{Reading:}

\begin{enumerate}

\item Chapter 7.7-7.8 of \textit{An Introduction to Computational Physics},
  by T. Pang (PDF; provided)

\item Chapter 10.1-10.5 of Zingale's \href{http://bender.astro.sunysb.edu/hydro_by_example/CompHydroTutorial.pdf}{Computational
  Hydrodynamics Tutorial} (PDF; provided in a previous pre-class
assignment).  Note: carefully read sections 10.1-10.3, and skim 10.4-10.5.

\end{enumerate}

\noindent
\textbf{Your assignment:}  

\begin{enumerate}

\item Exercise 10.2 in Zingale (p. 187): Implement a 1-D
\textit{explicit} diffusion solver for the domain [0,1] with Neumann
boundary conditions at each end and k = 1, using the discretization of
Eq. 10.6.  Begin with a Gaussian pulse as shown in Zingale's exercise
10.2, and compare to the analytic solution for solutions with $C =
0.8$ and $C = 2.0$ at the end point of the simulation.  What
difference do you see in the behavior of the solver in those two
cases?  When do you see numerical instabilities develop in either of
the cases, if it happens at all?  (Note: you may need to examine times
between $t=0$ and $t=t_{end}$ to see this.)

\item After doing the assigned reading, list at least two questions
that you have about solving diffusion-like equations (i.e., parabolic
PDEs) in the file \texttt{ANSWERS.md}.

\end{enumerate}

\vspace{5mm}

\noindent \textbf{Handing it in:} Include your code, your plots, and
your answers to the questions about (in the file \texttt{ANSWERS.md})
in your assignment.

\end{document}
