\documentclass[10pt]{article}
 \usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb,amsfonts,color, titling}
 \usepackage{listings}

\setlength{\droptitle}{-20mm} 

\usepackage[colorlinks=true,linkcolor=red,urlcolor=blue]{hyperref}

\title{Pre-class assignment \#15}
\author{PHY-905-005\\Computational Astrophysics and
  Astrostatistics\\Spring 2023}
 \date{} % leave blank to have no date

\begin{document}
 
\maketitle

\vspace{-10mm}

\noindent
\textbf{This assignment is due the evening of Wednesday March 22, 2023.}
 Turn in all materials via GitHub.

\vspace{5mm}

\noindent
\textbf{Reading:}

\begin{enumerate}

\item Chapters 2 and 3.1-3.7 of
  \href{http://astrostatistics.psu.edu/MSMA/}{Modern  Statistical
    Methods for Astronomy} by Feigelson and Babu.

\item Table of contents of the book \textit{All of Statistics: A
    Concise Course in Statistical Inference} by Larry Wasserman
  (Springer, 2004).  The PDF version is available through the MSU
  Library, and included in this pre-class assignment.  This is an
  incredibly useful resource for areas where Feigelson \& Babu's
  descriptions are a little bit too concise.  

\end{enumerate}

\noindent
Note that \textbf{we will be doing something a bit different in class}
-- I'll randomly select people to explain portions of the
reading to the entire class, so be prepared!

\vspace{5mm}

\noindent
\textbf{Your assignment:}  

\vspace{5mm}

\noindent
In the file \texttt{ANSWERS.md}, answer the following questions:

\begin{enumerate}

\item Why might we want to use Bayesian statistical analysis?

\item What is point estimation, and what are the relative strengths
  and weaknesses of the least-squares and maximum likelihood methods?

\item Why is resampling important in astronomy and astrophysics, and
  how do the jackknfe and bootstrap methods differ from each other? 

\item When attempting to determine whether a model is the ``best fit''
  to a given dataset, what are some of the choices you must
  make when deciding how to select your best-fit model?

\item After doing the assigned reading and answering the above
  questions, list at least three things that you are still puzzled by
  and would like to discuss more thoroughly in class.

\item Look through the table of contents of Feigelson \& Babu.  Which subjects
  seem like they will be the most useful for your research?  

\end{enumerate}

\end{document}
