\documentclass[10pt]{article}
 \usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb,amsfonts,color, titling,graphicx}
 \usepackage{listings}

\setlength{\droptitle}{-20mm} 

\usepackage[colorlinks=true,linkcolor=red,urlcolor=blue]{hyperref}

\title{In-class assignment \#14}
\author{PHY-905-005\\Computational Astrophysics and
  Astrostatistics\\Spring 2023}
 \date{} % leave blank to have no date

\begin{document}
 
\maketitle

\vspace{-5mm}


\noindent \textbf{Instructions:}   The file \texttt{functions.py}
includes two versions of the function

\begin{equation}
f(t) = 4 \sin(30 \pi t) + 1.5 \cos(60 \pi t) + 3 \cos(120 \pi t)
\end{equation}

\noindent
given to you in an array that spans the time range t$ = [0,4\pi]$~s.
By inspection, you can see that there are oscillations at frequencies
of 15, 30, and 60 Hz.  \textbf{Write a function} based on your
pre-class assignment that takes in the array of values of $f(t)$ (and
any other necessary information) and returns a new array containing
$g(t)$, which is $f(t)$ containing only oscillations that have
frequencies between $f_{min}$ and $f_{max}$, and verify that it is
behaving correctly by plotting both $f(t)$ and $g(t)$ together.  This
is what is known as a
\href{https://en.wikipedia.org/wiki/Band-pass_filter}{band-pass
filter}.  The special case with $f_{min} = 0, f_{max} \neq \infty$ is a low-pass filter,
and $f_{min}  \neq 0, f_{max} = \infty$ is a high-pass filter.  Answer the following
questions in the file \texttt{ANSWERS.md}:

\begin{enumerate}

\item How does the plot change as you increase/decrease the range
  $f_{min} - f_{max}$?  Why do you think that is?

\item Do you see any unexpected features?  What do you think is going
  on?

\item Try using the noisy version of the function.  Does this behave
  differently?  What happens as you increase the level of noise?

\item How might this type of filter be used in astronomical observations?

\end{enumerate}

\noindent
\textbf{Hand in} plots, code, and answers using the usual mechanism!

\end{document}