\documentclass[10pt]{article}
 \usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb,amsfonts,color, titling}
 \usepackage{listings}

\setlength{\droptitle}{-20mm} 

\usepackage[colorlinks=true,linkcolor=red,urlcolor=blue]{hyperref}

\title{In-class assignment \# 3}
\author{PHY-905-003\\Computational Astrophysics and
  Astrostatistics\\Spring 2023}
 \date{} % leave blank to have no date

\begin{document}
 
\maketitle

\vspace{-5mm}

\noindent \textbf{Instructions:} We're going to use the functions you
wrote for the Bisection and Newton's method, as well as some of the
additional root finding methods in the
\href{https://docs.scipy.org}{SciPy}
\href{https://docs.scipy.org/doc/scipy/reference/optimize.html#root-finding}{\texttt{optimize}}
package (specifically, the methods for scalar functions).

Your goal is straightforward: try to come up with multiple functions
f(x) that have one or more roots in an interval $[a,b]$ and which
break one or more of the following methods: the Bisection Method,
Newton's Method, the Secant Method (which is Newton's method with a
numerically-computed derivative -- use the function you wrote 
last time!), and
\href{http://mathworld.wolfram.com/BrentsMethod.html}{Brent's Method}.
In this case, ``break'' simply means to force the method to not find
an existing root in that interval within a finite number of iterations.

Experiment with several functions, intervals $[a,b]$, and all of the methods described
above.  What are the characteristics of functions or chosen intervals
$[a,b]$ that tend to break root-finding algorithms?  Are some
algorithms more robust than others to choice of function and/or interval?  Are any of these algorithms
effectively unbreakable?  Make some notes in \texttt{ANSWERS.md}, and
we'll also discuss it in class.

\vspace{5mm}

\noindent 
\textbf{What to turn in:} Turn in \texttt{ANSWERS.md}, any
source code you wrote, and any plots you created (and the scripts you used
to create them).  \textbf{Do not} turn in object files or
executables!

\end{document}